Bird beak's elaboration and innovation

### 2022-02-14 Analysis/09-Figure3.Rmd
Update 09-Figure3.Rmd

```{r, echo = FALSE, eval = FALSE}
## Getting the data to plot
global_data <- super_phylo_cor$phylogeny

@@

gg.density.plot(global_data)
```

@@

@@

    ## Plotting the results
    plot(cor.results$x, cor.results$y, pch = 19, xlab = ifelse(), ylab = ifelse(ID == 1 || ID == 5, "exploration", ""), main = cor.results$main, cex = pt.cex, col = col_vector)

@@

par(mfrow = c(2, 4))
lims <- list(x = range(global_data$x),
             y = range(global_data$y))

## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))

@@

for(i in 1:8) {
    plot.correlations(super_phylo_cor[[i]], xlim = lims$x, ylim = lims$x, tree = NULL, col.sub = col_subs[[i]], legend.pos = legends_pos[i])
}

```

>>>>

```{r, echo = FALSE, eval = FALSE}
library(ggplot2)
## Getting the data to plot
global_data <- super_phylo_cor$phylogeny

@@

pdf("../Manuscript/Figures/Figure3_b.pdf", height = 4, width = 4)
gg.density.plot(global_data)
dev.off()
```

Can't seem to integrate the density plot to the base R layout so I'll just do it in Inkscape for now

```{r, echo = FALSE, eval = FALSE}
pdf("../Manuscript/Figures/Figure3_a.pdf", height = 6, width = 3)
## Plotting the projection cartoon
library(mcmcmcglmmm)
## Get projection lines (from a centred matrix)
get.proj.line <- function(point, base) {
  point <- t(point)
  base <- t(base)
  proj_x <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[1];
  proj_y <- ( (point[1]*base[1] + point[2]*base[2]) / (base[1]*base[1] + base[2]*base[2]) ) * base[2];
  return(rbind(cbind(proj_x, proj_y), point))
}

set.seed(42)
## Space plot
cor_matrix <- matrix(cbind(1, 0.8, 0.8, 1), nrow = 2)
space <- space.maker(10, 2, rnorm, cor.matrix = cor_matrix)
lim <- c(floor(range(space)[1]), ceiling(range(space)[2]))

par(mfrow = c(2, 1))
## Projections per species
## Translate the major axis
par(bty = "n", mar = c(1,1,1,1)+0.1)
plot(space, pch = 19, xlab = "", ylab = "", xlim = lim, ylim = lim, col = "grey", main = "Panels B and C: species'\nprojection on the phylogeny", xaxt = "n", yaxt = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "grey", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "grey", lty = 1)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "orange")
# legend("bottomright", legend = "C")

## Do the projection
matrix <- space
point1 <- c(-1.2, 0)
point2 <- c(2, 0.75)
base_vector <- rbind(point1, point2)

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]

## Plot the projections
for(i in 1:nrow(matrix)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  lines(cent_proj, col = "darkgreen")
}

## Species projection onto their own axis
par(bty = "n", mar = c(1,1,1,1)+0.1)
plot(space, pch = 19, xlab = "", ylab = "", xlim = lim, ylim = lim, col = "grey", main = "Panels D: species'\nprojection on their own group", xaxt = "n", yaxt = "n")
abline(v = 0, col = "grey", lwd = 0.5)
abline(h = 0, col = "grey", lwd = 0.5)
lines(x = c(-1.2, 2), y = c(0, 0.75), lwd = 3, col = "grey")
## Plotting the ellipse
lines(ellipse::ellipse(cor_matrix, level = 0.75), col = "blue", lwd = 1)
## Plotting the major axis
lines(get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75), lwd = 3, col = "blue", lty = 1)
# legend("bottomright", legend = "D")

## Do the projection
matrix <- space
base_vector <- get.one.axis(list(VCV = cor_matrix, Sol = c(0,0)), level = 0.75)
point1 <- base_vector[1,]
point2 <- base_vector[2,]

## Get all the space (with the two last rows being the base vectors)
space_cent <- rbind(matrix, base_vector)
## Centre all the space
space_cent <- space_cent - rep(point1, rep.int(nrow(space_cent), ncol(space_cent)))
## Re-attribute the centred variables
matrix_cent <- space_cent[1:nrow(matrix), ]
base_vector_cent <- space_cent[-c(1:nrow(matrix)), ]
## Extend the base vector a little bit
extend_base <- base_vector
extend_base[1, ] <- extend_base[1,1] + 0.05
lines(extend_base, lwd = 1, col = "blue", lty = 2)

## Plot the projections
for(i in 1:nrow(matrix_cent)) {
  cent_proj <- get.proj.line(matrix_cent[i,], base_vector_cent[2, ])
  cent_proj[, 1] <- (cent_proj[, 1] + point1[1])
  cent_proj[, 2] <- (cent_proj[, 2] + point1[2])
  lines(cent_proj, col = "darkgreen")
}
dev.off()
```

@@

pdf("../Manuscript/Figures/Figure3_cd.pdf", height = 12, width = 8)
## Great the plot layout
layout_matrix <- cbind(matrix(0,2,2), matrix(0,2,2))
layout_matrix <- rbind(rbind(layout_matrix, matrix(1:8, 2, 4)), matrix(9:16, 2, 4))
plot_layout <- layout(layout_matrix)
# layout.show(plot_layout)

@@

    ## Plotting the results

    ## Selecting the x and y axis base on plot ID
    select.x <- function(ID, yes = "", no = "") {
        # ifelse(ID %in% c(5:8), yes, no)
        return(no)
    }
    select.y <- function(ID, yes = "", no = "") {
        # ifelse(ID == 1 || ID == 5, yes, no)
        return(no)
    }

    par(mar = c(select.x(ID, 4, 2), select.y(ID, 4, 2), 3, 1)+0.1)
    plot(cor.results$x, cor.results$y, pch = 19,
        xlab = select.x(ID, "elaboration"),
        ylab = select.y(ID, "explortation"),
        main = strsplit(cor.results$main, split = ":")[[1]][1],
        cex = pt.cex, col = col_vector,
        ...)

    if(!is.null(col.sub)) {
        legend_text <- levels(col_sub)
        if(length(empty <- which(legend_text == "")) > 0) {
            legend_text[empty] <- "other"
        }
        ## Adding the legend
        legend(x = legend.pos, legend = legend_text, pch = 19, cex = legend.cex, col = col_avail, bg = "white")
    }

@@

# par(mfrow = c(2, 4))
## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))

@@

for(i in 1:8) {
    one.plot.correlations(super_phylo_cor[[i]],col.sub = col_subs[[i]], legend.pos = legends_pos[i], ID = i, pt.cex = 0.8, legend.cex = 0.8)
}

## Fig 3c

# par(mfrow = c(2, 4))
## Getting the subsets for the colouring
col_subs <- c(split(shapespace$level2, f = shapespace$level1)[-1], "phylogeny" = list(shapespace$level1))

## Set the vector of legend positions
legends_pos <- c("bottomright", "topright", "topright", "topleft", "topleft", "topright", "bottomright", "bottomright")

for(i in 1:8) {
    one.plot.correlations(super_group_cor[[i]], col.sub = col_subs[[i]], legend.pos = legends_pos[i], ID = i, pt.cex = 0.8, legend.cex = 0.8)
}
dev.off()
```
